Supporting Information : The importance of a complete description of dependence between variables for ecology and related fields

### 2019-05-02 SupportingInformation.Rmd
minor edit
editing for length

---
title: "Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields"
author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance E. Loecke, Philip C. Reid, James D. Bever, Daniel C. Reuman"
geometry: "left=1cm,right=1cm,top=2.2cm,bottom=2.2cm"

@@

\textbf{Population data, plankton abundances.} 
The CPR plankton data presented in @Sheppard2017 was aggregated in the following way. Fifty 
$2^\circ$ by $2^\circ$ location boxes were examined between $-10^\circ$ and $10^\circ$ longitude 
and between $50^\circ$ and $60^\circ$ latitude (Fig. \ref{fig_plankton_map}). For each species, for each box, 
for each year, for each month, the average of any available CPR samples was taken to give monthly 
time series of population densities with some missing values. Missing monthly values were then 
replaced with the median value for that species for that location for that month. The data for each 
species for each box for each year was then averaged over all twelve months, and 26 usable annual 
time series were identified for wavelet analysis. The CPR plankton data used here is aggregated in 
the same way, except that for any species, box and year for which more than 4 months data were missing, 
missing values were not replaced with medians and those species, box, year combinations were not used.
We used only data from the 14 locations which had at least 45 years of annual data subject to this 
restriction, shown by blue markers in Fig. \ref{fig_plankton_map}. The CPR data are described in
great detail elsewhere, e.g., by  @Batten2003 and @Raitsos2014.

@@

To illustrate the potential for missing data to influence perceived
copula structure, we first generated 5000 points from a bivariate N copula
with parameter `r ncopparm` (Fig. \ref{fig:missingdata}A), and then 
randomly deleted `r 100*fracdelete`% of the points that occurred in the
upper-right triangle of the copula plot. Data were then re-ranked as
described in the Introduction to produce a new copula plot based
only on the censored dataset (Fig. \ref{fig:missingdata}B). Copula
structure in the tails was affected. Although real datasets have 
typically not been censored in the precise way we censored our
artificial data, it may be common for species characters to be 
measured on a subset of the extant species from the clade of interest,
and the probability of a species having been included in the dataset
may depend on one or both of the character values. 
For instance, if larger or smaller bird or mammal species were more likely
to have been measured for body mass or BMR, such accidental 
censoring of the data could have influenced our measurements of tail 
dependence. We have no reason to believe this is the case for our 
BMR versus body mass datasets, but we admit the possibility. Any systematic 
use of copulas to study life-history tradeoffs should take the possibility
into account of artefacts such as described here.

>>>>

---
title: "Supporting Information : The importance of a complete description of dependence between variables for ecology and related fields"
author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance E. Loecke, Philip C. Reid, James D. Bever, Daniel C. Reuman"
geometry: "left=1cm,right=1cm,top=2.2cm,bottom=2.2cm"

@@

\textbf{Environmental data, soil C and N.} Measurments were from the central pedon of a soil 
core, to 1m depth.

\textby{Phenological and population data on aphids.}
Data pre-processing for the aphid data was the same as 
that of @Sheppard2016,
but we excluded one of the 11 sites they used in order to only use sites with at least 30 
years data.

\textbf{Population data, plankton abundances.}
These data 
were obtained from the Continuous Plankton Recorder (CPR) dataset, now operated and maintained by 
the Marine Biological Association of the United Kingdom. Data pre-processing was similar but not identical to that 
of @Sheppard2017. 
The CPR plankton data presented in @Sheppard2017 was aggregated in the following way. Fifty 
$2^\circ$ by $2^\circ$ location boxes were examined between $-10^\circ$ and $10^\circ$ longitude 
and between $50^\circ$ and $60^\circ$ latitude (Fig. \ref{fig_plankton_map}). For each species, for each box, 
for each year, for each month, the average of any available CPR samples was taken to give monthly 
time series of population densities with some missing values. Missing monthly values were then 
replaced with the median value for that species for that location for that month. The data for each 
species for each box for each year was then averaged over all twelve months, and 26 usable annual 
time series were identified for wavelet analysis. The CPR plankton data used here is aggregated in 
the same way, except that for any species, box and year for which more than 4 months data were missing, 
missing values were not replaced with medians and those species, box, year combinations were not used.
We used only data from the 14 locations which had at least 45 years of annual data subject to this 
restriction, shown by blue markers in Fig. \ref{fig_plankton_map}. The CPR data are described in
great detail elsewhere, e.g., by  @Batten2003 and @Raitsos2014.

@@

To illustrate the potential for missing data to influence perceived
copula structure, we first generated 5000 points from a bivariate N copula
with parameter `r ncopparm` (Fig. \ref{fig:missingdata}A), and then 
randomly deleted `r 100*fracdelete`% of the points that occurred in the
upper-right triangle of the copula plot. Data were then re-ranked as
described in the \nameref{MT-Intro} to produce a new copula plot based
only on the censored dataset (Fig. \ref{fig:missingdata}B). Copula
structure in the tails was affected. Although real datasets have 
typically not been censored in the precise way we censored our
artificial data, it may be common for species characters to be 
measured on a subset of the extant species from the clade of interest,
and the probability of a species having been included in the dataset
may depend on one or both of the character values. 
For instance, if larger or smaller bird or mammal species were more likely
to have been measured for body mass or BMR, such accidental 
censoring of the data could have influenced our measurements of tail 
dependence. We have no reason to believe this is the case for our 
BMR versus body mass datasets, but we admit the possibility. Any systematic 
use of copulas to study life-history tradeoffs should take the possibility
into account of artefacts such as described here.
